Models

### 2024-11-13 analysis/05_model_metrics.Rmd
Attempts and mantel test

# Which predictor variables were significant?
slope_models %>% filter(term != "(Intercept)", p.value < 0.05) %>% View()
```

>>>>

# Which predictor variables were significant?
slope_models %>% filter(term != "(Intercept)", p.value < 0.05) %>% View()
```
Check if Jaccard distance is correlated with time with Mantel test
```{r}
library(vegan)

jaccard_dist <- model_data %>% 
    select(rarefyID, value, year, metric) %>% 
    pivot_wider(names_from = metric, values_from = value) %>% select(rarefyID, year, Jaccard_base) %>%
  unite(obsid, rarefyID, year, remove = FALSE) %>%
  rename(baselineID = rarefyID) %>%
  mutate(#baseline = is.na(Jaccard_base), 
         jacc_dist = 1-Jaccard_base) %>%
  group_by(baselineID) %>%
  mutate(time_dist = year - min(year)) %>%
  ungroup() %>%
  filter(!is.na(jacc_dist)) %>%
  select(-Jaccard_base, -year)

get_dist_mat <- function(df, dist_name){
  dist_row <- df %>% 
  select(obsid, baselineID, dist_name) %>% 
  pivot_wider(names_from = obsid, values_from = dist_name) %>% 
  mutate(baselineID = NA)

rownames(dist_row) <- "baselineID"

dist_mat <- dist_row %>%
  t() %>% as.data.frame()

dist_mat[,colnames(dist_row)[-1]] <- NA

return(as.dist(dist_mat))
}

# do it one by one for each rarefyID
get_mantel_stat <- function(rarefyID){
  dist_data <- jaccard_dist %>%
  filter(baselineID == rarefyID)

  jacc_dist_mat <- get_dist_mat(dist_data, "jacc_dist")
time_dist_mat <- get_dist_mat(dist_data, "time_dist")

mantel_test <- mantel(jacc_dist_mat, time_dist_mat, method = "spearman", permutations = 9999, na.rm = TRUE)

data.frame(rarefyID = rarefyID, p.value = mantel_test$signif)
}

mantel_stats <- purrr::map_dfr(unique(jaccard_dist$baselineID), get_mantel_stat)
```

```{r}
# create master matrices with all timeseries
master_jacc_dist <- purrr::map_dfr(unique(jaccard_dist$baselineID), function(x){
  #browser()
  dist_row <- jaccard_dist %>% 
    filter(baselineID == x) %>%
  select(obsid, baselineID, "jacc_dist") %>% 
  pivot_wider(names_from = obsid, values_from = "jacc_dist")

baseline_name <- unique(dist_row$baselineID)

dist_mat <- dist_row %>%
  mutate(baselineID = NA) %>%
  rename(!!baseline_name := baselineID) %>%
  t() %>% as.data.frame() %>%
  rename(!!baseline_name := V1)

dist_mat[,colnames(dist_row)[-1]] <- NA

return(dist_mat)
}) %>%
  as.dist()

master_time_dist <- purrr::map_dfr(unique(jaccard_dist$baselineID), function(x){
  #browser()
  dist_row <- jaccard_dist %>% 
    filter(baselineID == x) %>%
  select(obsid, baselineID, "time_dist") %>% 
  pivot_wider(names_from = obsid, values_from = "time_dist")

baseline_name <- unique(dist_row$baselineID)

dist_mat <- dist_row %>%
  mutate(baselineID = NA) %>%
  rename(!!baseline_name := baselineID) %>%
  t() %>% as.data.frame() %>%
  rename(!!baseline_name := V1)

dist_mat[,colnames(dist_row)[-1]] <- NA

return(dist_mat)
}) %>%
  as.dist()

mantel_test <- mantel(master_jacc_dist, master_time_dist, method = "spearman", permutations = 9999, na.rm = TRUE)
```
